Functions and Relations

### 2023-10-15 rec08/paper.tex
relations

\documentclass{amsart}
\input{decls}
\title{Functions}
\author{Frank Tsai}
\date{\today}
%\thanks{}
\begin{document}
\maketitle
\tableofcontents

@@

  The function $k$ has a unique code given by $\mathsf{encode}(k)$.
  Now, let's evaluate $k$ at its own code:
  \begin{align}
    k(\mathsf{encode}(k)) &= \mathsf{decode}(\mathsf{encode}(k))(\mathsf{encode}(k)) + 1\\
                          &= k(\mathsf{encode}(k)) + 1
  \end{align}
  This is a contradiction.\footnote{This proof is not a proof by contradiction. In fact, it is a constructively valid proof.}
\end{proof}

>>>>

\documentclass{amsart}
\input{decls}
\title{Functions and Relations}
\author{Frank Tsai}
\date{\today}
%\thanks{}
\begin{document}
\maketitle
\tableofcontents

\section{Relations}
\label{sec:relations}

\begin{defn}
  An $n$-ary \emph{relation} $R$ on a set $S$ is a subset:
  \[
    R \subseteq S^{n}
  \]
  We write $R(a,\ldots,z)$ whenever $(a,\ldots,z) \in R$.

  Binary relations will be the main focus of this class.
  For these relations, it is customary to use infix notations.
  That is, we write $aRb$ instead of $R(a,b)$.
\end{defn}

\begin{eg}
  The less-than-or-equal-to relation $\leq$ on $\dN$ is the subset
  \[
    \{(0,0), (0,1), \ldots, (1,1), (1,2), \ldots\} \subseteq \dN \times \dN
  \]
\end{eg}

\begin{eg}
  The divisibility relation $\mid$ on $\dZ$ is defined by
  \[
    a \mid b \iff \exists c.\,b = ac
  \]
  It is the subset
  \[
    \{(a,b) \in \dZ \times \dZ \mid \exists c.\,b = ac\}
  \]
\end{eg}

\begin{eg}
  The adjacency relation in a simple graph: two vertices $u$ are $v$ are adjacent if they are connected by an edge.
  It is the subset
  \[
    \{(u,v) \in V \times V \mid (u, v) \in E \vee (v, u) \in E\}
  \]
\end{eg}

\begin{defn}[Reflexivity]
  A binary relation $R$ on a set $S$ is \emph{reflexive} if for all elements $a$ of $S$, $aRa$.
  \[
    \forall a.\,aRa
  \]
\end{defn}

\begin{defn}[Symmetry]
  A binary relation $R$ on a set $S$ is \emph{symmetric} if for any two elements $a,b$ of $S$, if $aRb$ then $bRa$.
  \[
    \forall a.\,\forall b.\,aRb \imp bRa
  \]
\end{defn}

@@

  The function $k$ has a unique code given by $\mathsf{encode}(k)$.
  Now, let's evaluate $k$ at its own code:
  \begin{align}
    k(\mathsf{encode}(k)) &= \mathsf{decode}(\mathsf{encode}(k))(\mathsf{encode}(k)) + 1\\
                          &= k(\mathsf{encode}(k)) + 1
  \end{align}
  This is a contradiction.
\end{proof}
